---
title: 'Ch2: Statistical Learning Applied Exercises'
author: "Kylie"
date: "30/03/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # loads packages including ggplot2, dplyr
```

# Exercise 8

# Exercise 9

This exercise involves the Auto data set studied in the lab. Make sure
that the missing values have been removed from the data.

**(a) Which of the predictors are quantitative, and which are qualitative?**

```{r 9_a}
Auto <- read_csv("Auto.csv", na = c("", "NA", "?")) # loading data, making sure missing values (?s) are imported correctly
Auto <- na.omit(Auto) # removing missing values
sum(is.na(Auto)) # checking for missing values (this should be zero)

Auto <- mutate_if(Auto, is.character, factor) # converting characters to factors
Auto <- mutate(Auto, origin = as.factor(origin)) #converting origin to factor
```

Use `View(Auto)` to have a look at all of the data.

`mpg`, `displacement`, `horsepower`, `weight`, `acceleration`, and `year` are quantitative variables. `origin` and `name` are qualitative. `cylinder` appears to be quantiative, but only takes a few vaues (`r unique(as.factor(Auto$cylinders))`), so it may be better to treat this as a qualitative variable

**(b) What is the range of each quantitative predictor? You can answer this using the `range()` function.**

Including `cylinder` as quantitative for now.

```{r 9_b}

max_Auto <- summarise_if(Auto, is.numeric, list(max)) # calculating max for all numeric variables in Auto

min_Auto <- summarise_if(Auto, is.numeric, list(min)) # calculating min for all numeric variables in Auto

(range_Auto <- max_Auto - min_Auto) # evaluating and displaying range
```
**(c) What is the mean and standard deviation of each quantitative predictor?**

```{r 9_c}
# Mean:
(mean_Auto <- summarise_if(Auto, is.numeric, list(mean))) # calculating and displaying mean for all numeric variables in Auto
# Standard deviation:
(sd_Auto <- summarise_if(Auto, is.numeric, list(sd))) # calculating and displaying mean for all numeric variables in Auto
```

**(d) Now remove the 10th through 85th observations. What is the range, mean, and standard deviation of each predictor in the subset of the data that remains?**

```{r 9_d}
# Removing 10th to 85th observations:
Auto_short <- slice(Auto, -10:-85) # negatives indicate the rows should be dropped

# Range calculation:
max_Auto <- summarise_if(Auto_short, is.numeric, list(max)) # calculating max for all numeric variables in Auto
min_Auto <- summarise_if(Auto_short, is.numeric, list(min)) # calculating min for all numeric variables in Auto
(range_Auto <- max_Auto - min_Auto) # evaluating and displaying range

# Mean:
(mean_Auto <- summarise_if(Auto_short, is.numeric, list(mean))) # calculating and displaying mean for all numeric variables in Auto

# Standard deviation:
(sd_Auto <- summarise_if(Auto_short, is.numeric, list(sd))) # calculating and displaying mean for all numeric variables in Auto
```

**(e) Using the full data set, investigate the predictors graphically, using scatterplots or other tools of your choice. Create some plots highlighting the relationships among the predictors. Comment on your findings.**

**(f) Suppose that we wish to predict gas mileage (mpg) on the basis of the other variables. Do your plots suggest that any of the
other variables might be useful in predicting mpg? Justify your answer.**

# Exercise 10

# New packages/functions 

New packages/functions I've learnt during these exercises:

- `summarise_if`
- `slice`
