---
title: 'Ch2: Statistical Learning Applied Exercises'
author: "Kylie"
date: "30/03/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("GGally")
#install.packages("skimr")
library (MASS)
library(GGally) # for ggpairs
library(tidyverse) # loads packages including ggplot2, dplyr
library(skimr) # for nicer summaries using skim
library(cowplot) # for plot_grid
```

# Exercise 8

This exercise relates to the `College` data set, which can be found in the file `College.csv`. It contains a number of variables for 777 different universities and colleges in the US. The variables are:

- `Private`: Public/private indicator

- `Apps`: Number of applications received

- `Accept`: Number of applicants accepted

- `Enroll`: Number of new students enrolled

- `Top10perc`: New students from top 10% of high school class

- `Top25perc`: New students from top 25% of high school class

- `F.Undergrad`: Number of full-time undergraduates

- `P.Undergrad`: Number of part-time undergraduates

- `Outstate`: Out-of-state tuition

- `Room.Board`: Room and board costs

- `Books`: Estimated book costs

- `Personal`: Estimated personal spending

- `PhD`: Percent of faculty with Ph.D.’s

- `Terminal`: Percent of faculty with terminal degree

- `S.F.Ratio`: Student/faculty ratio

- `perc.alumni`: Percent of alumni who donate

- `Expend`: Instructional expenditure per student

- `Grad.Rate`: Graduation rate

**(a) Use the `read.csv()` function to read the data into R. Call the loaded data college. Make sure that you have the directory set
to the correct location for the data.**

```{r 8a}
# suppressing messages from readr
options(readr.num_columns = 0)

college <- read_csv('College.csv')
sum(is.na(college)) # checking for missing values (this should be zero)

```

**(b) Look at the data using the `fix()` function.** You should notice that the first column is just the name of each university.We don’t really want R to treat this as data. However, it may be handy to have these names for later. Try the following commands:

`rownames (college )=college [,1]`
`fix (college )`

You should see that there is now a row.names column with the
name of each university recorded. This means that R has given
each row a name corresponding to the appropriate university. R
will not try to perform calculations on the row names. However,
we still need to eliminate the first column in the data where the
names are stored. Try
`college =college [,-1]`
`fix (college )`

Now you should see that the first data column is Private. Note
that another column labeled row.names now appears before the
Private column. However, this is not a data column but rather
the name that R is giving to each row.

```{r 8b}
# suppressing messages from readr
options(readr.num_columns = 0)

college <- read_csv('College.csv')
sum(is.na(college)) # checking for missing values (this should be zero)

#fix(college)
college <- as.data.frame(college)
rownames(college) <- college[,1]
#fix (college)

college <- college[,-1]
#fix(college)
```

**(c) i. Use the `summary()` function to produce a numerical summary
of the variables in the data set.**

```{r 8_ci}
summary(college)

skim(college)
```

**ii. Use the `pairs()` function to produce a scatterplot matrix of
the first ten columns or variables of the data.** Recall that
you can reference the first ten columns of a matrix `A` using
`A[,1:10]`.

```{r 8_cii, fig.height = 19, fig.width = 15, message = FALSE}

pairs(select(college, 2:11))

# Plotting matrix of pair-wise scatterplots of first 10 variables
ggpairs(select(college, 2:11)) 
```


**iii. Use the `plot()` function to produce side-by-side boxplots of
Outstate versus Private.**

```{r 8_ciii}
ggplot(college, mapping = aes(x =Private, y = Outstate)) +
  geom_boxplot()
```


**iv. Create a new qualitative variable, called `Elite`, by binning
the `Top10perc variable`.** We are going to divide universities
into two groups based on whether or not the proportion
of students coming from the top 10% of their high school
classes exceeds 50%.

`Elite =rep ("No",nrow(college ))`

`Elite [college$Top10perc >50]=" Yes"`

`Elite =as.factor (Elite)`

`college =data.frame(college ,Elite)`

Use the `summary()` function to see how many elite universities
there are. Now use the `plot()` function to produce
side-by-side boxplots of `Outstate` versus `Elite`.

```{r 8civ}

college <- mutate(college, Elite = as.factor(case_when(Top10perc > 50 ~ "Yes", 
                             Top10perc <= 50 ~ "No")))

skim(select(college, Elite))

ggplot(college, mapping = aes(x = Elite, y = Outstate)) +
  geom_boxplot()
```

**v. Use the `hist()` function to produce some histograms with
differing numbers of bins for a few of the quantitative variables.**
You may find the command `par(mfrow=c(2,2))` useful:
it will divide the print window into four regions so that four
plots can be made simultaneously. Modifying the arguments
to this function will divide the screen in other ways.

```{r 8cv}
# trying bin width of 100
A <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Apps), binwidth = 100)

B <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Accept), binwidth = 100)

C <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Enroll), binwidth = 100)

D <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Top10perc), binwidth = 100)

plot_grid(A, B, C, D)

# trying bin width of 10
A <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Apps), binwidth = 10)

B <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Accept), binwidth = 10)

C <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Enroll), binwidth = 10)

D <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Top10perc), binwidth = 10)

plot_grid(A, B, C, D)

# trying bin width of 1
A <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Apps), binwidth = 1)

B <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Accept), binwidth = 1)

C <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Enroll), binwidth = 1)

D <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Top10perc), binwidth = 1)

plot_grid(A, B, C, D)

# trying a mix of bin widths
A <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Apps), binwidth = 1000)

B <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Accept), binwidth = 1000)

C <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Enroll), binwidth = 100)

D <- ggplot(data = college) +
  geom_histogram(mapping = aes(x = Top10perc), binwidth = 1)

plot_grid(A, B, C, D)

```

**vi. Continue exploring the data, and provide a brief summary
of what you discover.**

# Exercise 9

This exercise involves the `Auto` data set studied in the lab. Make sure that the missing values have been removed from the data.

**(a) Which of the predictors are quantitative, and which are qualitative?**

```{r 9_a}
# suppressing messages from readr
options(readr.num_columns = 0)

Auto <- read_csv("Auto.csv", na = c("", "NA", "?")) # loading data, making sure missing values (?s) are imported correctly
Auto <- na.omit(Auto) # removing missing values
sum(is.na(Auto)) # checking for missing values (this should be zero)

Auto <- mutate_if(Auto, is.character, factor) # converting characters to factors
Auto <- mutate(Auto, origin = as.factor(origin)) #converting origin to factor
```

Use `View(Auto)` to look at all of the data.

`mpg`, `displacement`, `horsepower`, `weight`, `acceleration`, and `year` are quantitative variables. `origin` and `name` are qualitative. `cylinder` appears to be quantiative, but only takes a few vaues (`r unique(as.factor(Auto$cylinders))`), so it may be better to treat this as a qualitative variable

**(b) What is the range of each quantitative predictor? You can answer this using the `range()` function.**

Including `cylinder` as quantitative for now.

```{r 9_b}

max_Auto <- summarise_if(Auto, is.numeric, list(max)) # calculating max for all numeric variables in Auto

min_Auto <- summarise_if(Auto, is.numeric, list(min)) # calculating min for all numeric variables in Auto

(range_Auto <- max_Auto - min_Auto) # evaluating and displaying range
```
**(c) What is the mean and standard deviation of each quantitative predictor?**

```{r 9_c}
# Mean:
(mean_Auto <- summarise_if(Auto, is.numeric, list(mean))) # calculating and displaying mean for all numeric variables in Auto
# Standard deviation:
(sd_Auto <- summarise_if(Auto, is.numeric, list(sd))) # calculating and displaying mean for all numeric variables in Auto
```

**(d) Now remove the 10th through 85th observations. What is the range, mean, and standard deviation of each predictor in the subset of the data that remains?**

```{r 9_d}
# Removing 10th to 85th observations:
Auto_short <- slice(Auto, -10:-85) # negatives indicate the rows should be dropped

# Range calculation:
max_Auto <- summarise_if(Auto_short, is.numeric, list(max)) # calculating max for all numeric variables in Auto
min_Auto <- summarise_if(Auto_short, is.numeric, list(min)) # calculating min for all numeric variables in Auto
(range_Auto <- max_Auto - min_Auto) # evaluating and displaying range

# Mean:
(mean_Auto <- summarise_if(Auto_short, is.numeric, list(mean))) # calculating and displaying mean for all numeric variables in Auto

# Standard deviation:
(sd_Auto <- summarise_if(Auto_short, is.numeric, list(sd))) # calculating and displaying mean for all numeric variables in Auto
```

**(e) Using the full data set, investigate the predictors graphically, using scatterplots or other tools of your choice. Create some plots highlighting the relationships among the predictors. Comment on your findings.**

```{r 9_e, fig.height = 19, fig.width = 15, message = FALSE}
# suppressing messages from readr
options(readr.num_columns = 0)

Auto <- read_csv("Auto.csv", na = c("", "NA", "?")) # loading data, making sure missing values (?s) are imported correctly
Auto <- na.omit(Auto) # removing missing values
Auto <- mutate_if(Auto, is.character, factor) # converting characters to factors
Auto <- mutate(Auto, origin = as.factor(origin)) #converting origin to factor

# Plotting matrix of pair-wise scatterplots of all variables
ggpairs(select(Auto, -name)) # not including name because it has too many categories

ggplot(data = Auto) +
  geom_bar(mapping = aes(x = name)) # bar chart of name. Not very useful because there are so many categories
```

Key points:

- Many of the variables are strongly correlated. For example, `weight` has a correlation coefficient with an absolute value higher than 0.8 with `mpg`, `cylinders`, `displacement` and `horsepower`. 
- The distribution of `mpg` is slightly negatively skewed.
- There are some linear relationships between some variables. e.g. between: `cylinders` and `displacement`, `displacement` and `horsepower`, `displacement` and `weight`, `horsepower` and `weight`.
- Relationships between `year` and the other variables are very noisy, although to a lesser extent for `mpg` and `acceleration`.
- `origin` affects `mpg` (e.g. higher `mpg` for `origin = 3`), but that could be because of changes in `origin` over time (because `year` also affects `mpg`) rather than a direct effect of `origin` on `mpg`.

**(f) Suppose that we wish to predict gas mileage (mpg) on the basis of the other variables. Do your plots suggest that any of the other variables might be useful in predicting mpg? Justify your answer.**

```{r 9_f, fig.height = 12, fig.width = 15, message = FALSE}
# suppressing messages from readr
options(readr.num_columns = 0)

Auto <- read_csv("Auto.csv", na = c("", "NA", "?")) # loading data, making sure missing values (?s) are imported correctly
Auto <- na.omit(Auto) # removing missing values
# Not converting characters and origin to factors because this causes problems with `gather`

Auto_wrap <- gather(select(Auto, -name), -mpg, key = "vars", value = "values") # getting data in the right format to use for facet_wrap. Excluding `name` because including it messes up the x-axes for facet_wrap 

  ggplot(Auto_wrap, aes(x = values, y = mpg)) +
    geom_point() +
    facet_wrap(~ vars, scales = "free_x", ncol = 2)

```

All of the variables except `name` look like they would be useful for predicting, with `displacement`, `horsepower`, `weight` and `cylinder` looking particularly useful because of their strong correlations with `mpg` (see above figure and correlation coefficients in 9(e)). `name` is not very useful (no clear relationship with `mpg` and it would introduce way too many dummy variables).

# Exercise 10

**(a) To begin, load in the `Boston` data set. The Boston data set is part of the `MASS` library in R. How many rows are in this data set? How many columns? What do the rows and columns represent?**

```{r 10a}
library (MASS)
# Now the data set is contained in the object Boston.
as_tibble(Boston)

#Read about the data set:
#?Boston
```

The data set has 506 rows and 14 columns. The rows represent towns. The columns represent the following variables:

 - `crim`: per capita crime rate by town.
 - `zn`: proportion of residential land zoned for lots over 25,000 sq.ft.
 - `indus`: proportion of non-retail business acres per town.
 - `chas`: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
 - `nox`: nitrogen oxides concentration (parts per 10 million).
 - `rm`: average number of rooms per dwelling.
 - `age`: proportion of owner-occupied units built prior to 1940.
 - `dis`: weighted mean of distances to five Boston employment centres.
 - `rad`: index of accessibility to radial highways.
 - `tax`: full-value property-tax rate per \$10,000.
 - `ptratio`: pupil-teacher ratio by town.
 - `black`: 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.
 - `lstat`: lower status of the population (percent).
 - `medv`: median value of owner-occupied homes in \$1000s.

**(b) Make some pairwise scatterplots of the predictors (columns) in this data set. Describe your findings.**

```{r 10_b, fig.height = 20, fig.width = 15, message = FALSE}
Boston_10b <- as_tibble(Boston) %>%
  mutate_if(is.integer, factor) # converting integers to factors

# Plotting matrix of pair-wise scatterplots of all variables
ggpairs(Boston_10b) 

```

**(c) Are any of the predictors associated with per capita crime rate? If so, explain the relationship.**

```{r 10_c, fig.height = 12, fig.width = 15}
Boston_wrap <- as_tibble(Boston) %>%
  gather(-crim, key = "vars", value = "values") # getting data in the right format to use for facet_wrap. Excluding `name` because including it messes up the x-axes for facet_wrap 

  ggplot(Boston_wrap, aes(x = values, y = crim)) +
    geom_point() +
    facet_wrap(~ vars, scales = "free_x", ncol = 3)

```

**(d) Do any of the suburbs of Boston appear to have particularly high crime rates? Tax rates? Pupil-teacher ratios? Comment on the range of each predictor.**

```{r 10_d, message = FALSE}
Boston_10d <- as_tibble(Boston) %>%
  mutate_if(is.integer, factor)# converting integers to factors

ggplot(Boston_10d) + 
  geom_histogram(mapping = aes(x = crim))

ggplot(Boston_10d, mapping = aes(y = crim)) +
  geom_boxplot()

ggplot(Boston_10d) + 
  geom_histogram(mapping = aes(x = tax))

ggplot(Boston_10d, mapping = aes(y = tax)) +
  geom_boxplot()

ggplot(Boston_10d) + 
  geom_histogram(mapping = aes(x = ptratio))

ggplot(Boston_10d, mapping = aes(y = ptratio)) +
  geom_boxplot()

skim(Boston_10d)
```


**(e) How many of the suburbs in this data set bound the Charles river?**

`r sum(Boston$chas)` suburbs bound the Charles River.

**(f) What is the median pupil-teacher ratio among the towns in this data set?**

Median = `r median(Boston$ptratio)`

**(g) Which suburb of Boston has lowest median value of owneroccupied homes? What are the values of the other predictors for that suburb, and how do those values compare to the overall ranges for those predictors? Comment on your findings.**

```{r 10_g}
# suburb of Boston has lowest median value of owneroccupied homes
which(Boston$medv == min(Boston$medv))

# values of predictors for suburbs with the lowest median value of owneroccupied homes
kable(filter(Boston, medv == min(medv)))

# Range calculation:
kable(max_Bost <- summarise_if(Boston, is.numeric, list(max))) # calculating max for all numeric variables in Boston
kable(min_Bost <- summarise_if(Boston, is.numeric, list(min))) # calculating min for all numeric variables in Boston

```


**(h) In this data set, how many of the suburbs average more than seven rooms per dwelling? More than eight rooms per dwelling? Comment on the suburbs that average more than eight rooms per dwelling.**

```{r 10_h}

# suburbs average more than seven rooms per dwelling
filter(Boston, rm > 7) %>%
  count()

# suburbs average more than eight rooms per dwelling
filter(Boston, rm > 8) %>%
  count()

kable(filter(Boston, rm > 8))
```

# New packages/functions 

New packages/functions I've learnt during these exercises:

- `summarise_if`
- `slice`
- `gather`
- `case_when`
- use `as_tibble()` instead of `as.tibble()`