---
title: 'Ch3: Linear Regression Applied Exercises'
author: "Kylie"
date: "30/03/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("GGally")
#install.packages("skimr")
library(ISLR) # data sets for the ISLR book
library (MASS)
library(GGally) # for ggpairs
library(tidyverse) # loads packages including ggplot2, dplyr
library(skimr) # for nicer summaries using skim
library(cowplot) # for plot_grid
```

# Exercise 8

This question involves the use of simple linear regression on the `Auto`
data set.

**(a) Use the `lm()` function to perform a simple linear regression with `mpg` as the response and `horsepower` as the predictor.** Use the `summary()` function to print the results. Comment on the output.

```{r 8_a}

lm_Auto <- lm(mpg ~ horsepower, data = Auto)

summary(lm_Auto)
```

For example:

**i. Is there a relationship between the predictor and the response?**

Yes there is a statistically significant relationship between the predictor and the response.

**ii. How strong is the relationship between the predictor and the response?**

Does this refer to the statistical significance or the size of the effect?

A one unit change in horsepower causes a 0.15 unit decrease in miles per gallon. Since the range of `mpg` is `r range(Auto$mpg)` then this doesn't seem like a large/strong effect.

**iii. Is the relationship between the predictor and the response positive or negative?**

Negative.

**iv. What is the predicted `mpg` associated with a `horsepower` of 98? What are the associated 95% confidence and prediction intervals?**

```{r 8_a_iv}

lm_Auto <- lm(mpg ~ horsepower, data = Auto)

summary(lm_Auto)
```

**(b) Plot the response and the predictor. Use the `abline()` function to display the least squares regression line.**

**(c) Use the `plot()` function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.**

# New packages/functions 

New packages/functions I've learnt during these exercises:

