---
title: "Useful Functions"
author: "Kylie Foster"
date: "1 May 2019"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(GGally) # for ggpairs
library(tidyverse) # loads packages including ggplot2
```

---

# ISL Chapter 2

### `ggpairs()`

Useful for exploratory data analysis. Produces a pairwise comparison of multivariate data. By default, `ggpairs` provides two different comparisons of each pair of columns and displays either the density (for continous variables) or count (for categorical variables) along the diagonal.

To carry out a pairwise comparison of all variables in a dataframe:

```{r ggpairs_all, message=FALSE}
data(tips, package = "reshape")
(pm <- ggpairs(tips))
```

If this creates too many subplots you can restrict the plot to a specific number of columns using:

```{r ggpairs_restrict, message=FALSE}

(pm <- ggpairs(tips, columns = c("total_bill", "time", "tip")))
# or equivalently you can use column numbers instead of variable names:
pm <- ggpairs(tips, columns = c(1, 6, 2))
```

To change the labels on the axes use `columnLabels`:

```{r ggpairs_labels, message=FALSE}
(pm <- ggpairs(tips, columns = c("total_bill", "time", "tip"), columnLabels = c("Total Bill", "Time of Day", "Tip")))

```

To apply aesthetics to all of the subplots (e.g. colour) use the `mapping` parameter:

```{r ggpairs_color, message=FALSE}
(pm <- ggpairs(tips, mapping = aes(color = sex), columns = c("total_bill", "time", "tip")))
```

It is also possible to change the `theme`, for example:

```{r ggpairs_theme, message=FALSE}
(pm <- ggpairs(tips, mapping = aes(color = sex), columns = c("total_bill", "time", "tip")) +
   theme_bw())
```

For more information (including how to replace default plots with different types of plots, add trendlines etc):
https://mran.microsoft.com/snapshot/2016-01-07/web/packages/GGally/vignettes/ggpairs.html

### `glimpse()`

This is like a transposed version of print: columns run down the page, and data runs across. This is more useful than the default display of tibbles because it makes it possible to see every column in a data frame/tibble. The width of the displayed data defaults to the width of the screen.

Example:

```{r}
glimpse(mtcars)
```

Source: Rdocumentation

- `summarise_if`
- `slice`
- `gather` sometimes required to use facet_wrap
- `case_when`

### Other notes:

- To suppress messages from readr (when using `read_csv`) in Rmarkdown output, use `options(readr.num_columns = 0)`.

- Use `message = FALSE` to prevent Rmarkdown output from displaying messages such as "`stat_bin()` using `bins = 30`. Pick better value with `binwidth`."

- One way to check for missing values is to use: `sum(is.na(data))`. This will return the number of missing values.

- You can use `fig.height =` and `fig.width =` in the heading of a code chunk to change the dimensions of all figures in that chunk in the Rmarkdown output.

- use `as_tibble()` instead of `as.tibble()`

# ISL Chapter 4

- `glm()`

- `predict()`

- `lda()`. Used to fit a Linear Discriminant Analysis model. Part of the `MASS` library. The syntax is identical to that of `lm()`. LDA assumes that the variables are continuous, so it is not a good idea to use it when you have categorical variables (use logistic regression instead).

- `qda()`. Used to fit a Quadratic Discriminant Analysis model. Part of the `MASS` library. Syntax is identical to that of `lda()`. Again, discriminant analysis assumes the variables are continous, so it is not a good idea to use it when you have categorical variables.

## `knn()`. 
Used to fit K-nearest neighbors. Part of the `class` library. This function works differently to the other model-fitting
functions described above. Instead of a two-step
approach in which we first fit the model and then we use the model to make
predictions, `knn()` fits the model and forms predictions using a single command. The function requires four inputs.

1. A matrix containing the predictors associated with the **training data**.
2. A matrix containing the predictors associated with the data for which
we wish to make predictions (the **testing data**).
3. A vector containing the **class labels for the training observations**.
4. A value for **K, the number of nearest neighbors** to be used by the
classifier.

- `sample()`