---
title: "An Introduction to Statistical Learning: Chapter 5 Applied Exercises"
author: "Kylie Foster"
date: "8 May 2019"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS) # for lda, and contains a large collection of datasets
library(tidyverse) # loads packages including ggplot2, dplyr
library(ISLR) # data sets for the ISLR book
library(boot) # for cv.glm() and boot()
```

## Exercise 5

In Chapter 4, we used logistic regression to predict the probability of
default using `income` and `balance` on the `Default` data set. We will
now estimate the test error of this logistic regression model using the
validation set approach. Do not forget to set a random seed before
beginning your analysis.

#### (a) Fit a logistic regression model that uses income and balance to predict default.

```{r 5_a}

log_fit <- glm(default ~ income + balance, data = Default, family = "binomial")
summary(log_fit)
```

#### (b) Using the validation set approach, estimate the test error of this model. In order to do this, you must perform the following steps:

#### i. Split the sample set into a training set and a validation set.

```{r 5bi}

set.seed(1)
# vector of row numbers to using in training set.
train <- sample(dim(Default)[1], dim(Default)[1]*0.8, replace = FALSE, prob = NULL) # using dim[1] 
# to work out the correct length and using 80% of the data for training

```

#### ii. Fit a multiple logistic regression model using only the training observations.

```{r 5bii}
# fitting logistic model using only training data
log_fit_train <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)

```

#### iii. Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5.

```{r 5biii}
# calculating predictions using test set
log_prob <- predict(log_fit_train, Default, type = "response")[-train] # predicted probabilities for the test data

# Getting class predictions using 0.5 as the threshold:
log_pred <- as.factor(case_when(log_prob <= 0.5 ~"No",
                                log_prob > 0.5 ~ "Yes"))

```

#### iv. Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.

```{r 5biv}

mean(log_pred != Default$default[-train]) # validation set error

```

#### (c) Repeat the process in (b) three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained.

```{r}
# Using a different seed to give a different train/test split
set.seed(2)
train2 <- sample(dim(Default)[1], 0.8*dim(Default)[1])
# fitting logistic model using train2
log_fit_train2 <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train2)
# calculating predictions using test set
log_prob <- predict(log_fit_train2, Default, type = "response")[-train2]
log_pred <- as.factor(case_when(log_prob <= 0.5 ~ "No",
                                log_prob > 0.5 ~ "Yes"))
# validation test set error
mean(log_pred != Default$default[-train2])

##
# Using another different seed to give a different train/test split
set.seed(3)
train3 <- sample(dim(Default)[1], 0.8*dim(Default)[1])
# fitting logistic model using train3
log_fit_train3 <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train3)
# calculating predictions using test set
log_prob <- predict(log_fit_train3, Default, type = "response")[-train3]
log_pred <- as.factor(case_when(log_prob <= 0.5 ~ "No",
                                log_prob > 0.5 ~ "Yes"))
# validation test set error
mean(log_pred != Default$default[-train3])

```

The validation set errors are similar, low, values for the different splits (approximately 2 to 3%).

#### (d) Now consider a logistic regression model that predicts the probability of `default` using `income`, `balance`, and a dummy variable for `student`. 
Estimate the test error for this model using the validation set approach. Comment on whether or not including a
dummy variable for `student` leads to a reduction in the test error rate.

```{r 5d}

set.seed(1)
train <- sample(dim(Default)[1], 0.8*dim(Default)[1]) # 80%/20% train/test split
# gitting logistic model, including student as a predictor
log_fit_train <- glm(default ~., data = Default, family = "binomial", subset = train)
summary(log_fit_train)
# calculating predicted probabilities using test (cross validation) set
log_prob <- predict(log_fit_train, Default, type = "response")[-train]
log_pred <- as.factor(case_when(log_prob <= 0.5 ~ "No",
                                log_prob > 0.5 ~ "Yes"))
# validation test set error
mean(log_pred != Default$default[-train])

```

Using this validation set, including a dummy variable for `student` did not lead to any meaningful reducion in the test error rate.

## Exercise 6

We continue to consider the use of a logistic regression model to
predict the probability of `default` using `income` and `balance` on the
`Default` data set. In particular, we will now compute estimates for
the standard errors of the `income` and `balance` logistic regression coefficients
in two different ways: (1) using the bootstrap, and (2) using
the standard formula for computing the standard errors in the `glm()`
function. Do not forget to set a random seed before beginning your
analysis.

#### (a) Using the `summary()` and `glm()` functions, determine the estimated standard errors for the coefficients associated with `income` and `balance` in a multiple logistic regression model that uses both predictors.

```{r 6a}
# fitting multiple logistic regression model
log_multi <- glm(default ~ income + balance, data = Default, family = "binomial")
summary(log_multi) # using summary to determine the estimated standard errors for the coefficients

```

#### (b) Write a function, `boot.fn()`, that takes as input the `Default` data set as well as an index of the observations, and that outputs the coefficient estimates for `income` and `balance` in the multiple logistic regression model.

```{r 6b}

boot_fn <- function(data, index){
  coef(glm(default ~ income + balance, data = data, family = "binomial", subset = index))
}

```


#### (c) Use the `boot()` function together with your `boot.fn()` function to estimate the standard errors of the logistic regression coefficients for `income` and `balance`.

```{r 6c}
set.seed(1)
boot(Default, boot_fn, 1000) # calculating 1000 bootstrap estimates
```


#### (d) Comment on the estimated standard errors obtained using the `glm()` function and using your bootstrap function.

The standard errors obtained using the two methods are very similar.

## Exercise 7

In Sections 5.3.2 and 5.3.3, we saw that the `cv.glm()` function can be
used in order to compute the LOOCV test error estimate. Alternatively,
one could compute those quantities using just the `glm()` and `predict.glm()` functions, and a for loop. You will now take this approach
in order to compute the LOOCV error for a simple logistic
regression model on the `Weekly` data set. Recall that in the context
of classification problems, the LOOCV error is given in (5.4).

#### (a) Fit a logistic regression model that predicts `Direction` using `Lag1` and `Lag2`.

```{r 7a}
# fitting multiple logistic regression model
log_Weekly <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = "binomial")
```

#### (b) Fit a logistic regression model that predicts `Direction` using `Lag1` and `Lag2` using all but the first observation.

```{r 7b}

index <- 2:dim(Weekly)[1]

log_Weekly_minus1 <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = "binomial", subset = index)
```

#### (c) Use the model from (b) to predict the direction of the first observation. You can do this by predicting that the first observation will go up if P(Direction="Up"|Lag1, Lag2) > 0.5. Was this observation correctly classified?

```{r 7c}
# calculating probability of Up for first observation
log_prob <- predict(log_Weekly_minus1, Weekly[1, ], type = "response")
# classifying first observation as Up of Down
log_pred <- as.factor(case_when(log_prob > 0.5 ~ "Up",
                                log_prob <= 0.5 ~ "Down"))
# checking if this observation was correctly classified
case_when(log_pred == Weekly$Direction[1] ~ "Correctly classified",
          log_pred != Weekly$Direction[1] ~ "Incorrectly classified")

```


#### (d) Write a for loop from i = 1 to i = n, where n is the number of observations in the data set, that performs each of the following steps:

#### i. Fit a logistic regression model using all but the ith observation to predict Direction using Lag1 and Lag2.

#### ii. Compute the posterior probability of the market moving up for the ith observation.

#### iii. Use the posterior probability for the ith observation in order to predict whether or not the market moves up.

#### iv. Determine whether or not an error was made in predicting the direction for the ith observation. If an error was made, then indicate this as a 1, and otherwise indicate it as a 0.

```{r 7d}
n <- dim(Weekly)[1] # number of observations in the data set
err <- rep(0, n) # initialising vector
for (i in 1:n) {
  index <- (1:n)[-i] # removing row i from index
  # fitting logistic model for all observations except i
  log_fit <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = "binomial", subset = index)  
  # finding the probability of Up for observation i
  log_prob <-  predict(log_fit, Weekly[i, ], type = "response")
  # converting the above probability to a direction
  log_pred <- as.factor(case_when(log_prob > 0.5 ~ "Up",
                                  log_prob < 0.5 ~ "Down"))
  # determining if an error was made predicting the ith observation, and saving the results
  err[i] <- case_when(log_pred == Weekly$Direction[i] ~ 0, # correct
          log_pred != Weekly$Direction[i] ~ 1) # incorrect
}

```

#### (e) Take the average of the n numbers obtained in (d)iv in order to obtain the LOOCV estimate for the test error. Comment on the results.

```{r 7e}

mean(err) # LOOCV estimate of the test error

```
The error rate is quite high (approx 45%).

## Exercise 8

We will now perform cross-validation on a simulated data set.

#### (a) Generate a simulated data set as follows:

`set.seed(1)`

`x = rnorm(100)`

`y = x - 2*x^2 + rnorm(100)`

In this data set, what is n and what is p? Write out the model
used to generate the data in equation form.

```{r 8a}
set.seed(1)
x <- rnorm(100)
y <- x - 2*x^2 + rnorm(100)

```

The number of observations is 100 (n = 100) and the number of predictors is 2, `x` and `x^2` (p = 2).

$y = \beta_1x + \beta_2x^2 + \epsilon$, with $\beta_1 = 1$ and $\beta_2 = -2$.

#### (b) Create a scatterplot of `X` against `Y`. Comment on what you find.

```{r 8b}

plot(x, y)

```

There is (unsurprisingly) a quadratic relationship between `x` and `y` with some noise present.

#### (c) Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using least squares:

#### i. $Y = \beta_0 + \beta_1X + \epsilon$

```{r 8ci}
set.seed(1)

data_xy <- data.frame(x, y) # combining x and y into a single dataframe

glm_fit <- glm(y ~ x, data = data_xy) # fitting eq for 8ci
summary(glm_fit)

cv.glm(data_xy, glm_fit)$delta[1] # LOOCV error
```


#### ii. $Y = \beta_0 + \beta_1X + \beta_2X^2 + \epsilon$

```{r 8cii}

glm_fit <- glm(y ~ x + I(x^2), data = data_xy) # fitting eq for 8cii
summary(glm_fit)

cv.glm(data_xy, glm_fit)$delta[1] # LOOCV error
```


#### iii. $Y = \beta_0 + \beta_1X + \beta_2X^2 +\beta_3X^3 + \epsilon$

```{r 8ciii}

glm_fit <- glm(y ~ x + I(x^2) + I(x^3), data = data_xy) # fitting eq for 8ciii
summary(glm_fit)

cv.glm(data_xy, glm_fit)$delta[1] # LOOCV error
```


#### iv. $Y = \beta_0 + \beta_1X + \beta_2X^2 + \beta_3X^3 + \beta_4X^4 + \epsilon$.

```{r 8civ}

glm_fit <- glm(y ~ x + I(x^2) + I(x^3) + I(x^4), data = data_xy) # fitting eq for 8civ
summary(glm_fit)

cv.glm(data_xy, glm_fit)$delta[1] # LOOCV error
```

Note you may find it helpful to use the data.frame() function
to create a single data set containing both `X` and `Y`.

#### (d) Repeat (c) using another random seed, and report your results. Are your results the same as what you got in (c)? Why?

```{r 8d}
set.seed(10)

glm_fit <- glm(y ~ x, data = data_xy) # fitting eq for 8ci
cv.glm(data_xy, glm_fit)$delta[1] # LOOCV error

glm_fit <- glm(y ~ x + I(x^2), data = data_xy) # fitting eq for 8cii
cv.glm(data_xy, glm_fit)$delta[1] # LOOCV error

glm_fit <- glm(y ~ x + I(x^2) + I(x^3), data = data_xy) # fitting eq for 8ciii
cv.glm(data_xy, glm_fit)$delta[1] # LOOCV error

glm_fit <- glm(y ~ x + I(x^2) + I(x^3) + I(x^4), data = data_xy) # fitting eq for 8civ
cv.glm(data_xy, glm_fit)$delta[1] # LOOCV error
```

The results of LOOCV are identical for different random seeds (unless the set.seed() used before calculating x and y are changed), this is because all observations are used to calculate the LOOCV error (there is no random sample of values used).

#### (e) Which of the models in (c) had the smallest LOOCV error? Is this what you expected? Explain your answer.

The model in **8(c)(ii)** ($Y = \beta_0 + \beta_1X + \beta_2X^2 + \epsilon$) has the smallest LOOCV error. This is as expected because the data was simulated using a quadratic function.

#### (f) Comment on the statistical significance of the coefficient estimates that results from fitting each of the models in (c) using least squares. Do these results agree with the conclusions drawn based on the cross-validation results?

In models (c)(ii) to (c)(iv) only the coefficients of `x` and `x^2` are statistically significant, the intercept term and coefficients of `x` terms with a power higher than 2 are not statistically significant. This is expected because the data was simulated using an equation with only `x` and `x^2` terms (with no intercept and no higher power terms).

## Exercise 9

We will now consider the `Boston` housing data set, from the `MASS` library.

#### (a) Based on this data set, provide an estimate for the population mean of `medv`. Call this estimate $\hat{\mu}$.

```{r 9a}
(mu_hat <- mean(Boston$medv)) # estimate for the population mean of medv
```


#### (b) Provide an estimate of the standard error of $\hat{\mu}$. Interpret this result.

Hint: We can compute the standard error of the sample mean by
dividing the sample standard deviation by the square root of the
number of observations.

```{r 9b}
(mu_hat_SE <- sd(Boston$medv)/sqrt(length(Boston$medv)))
```

Not sure how to interpret this result.

#### (c) Now estimate the standard error of $\hat{\mu}$ using the bootstrap. How does this compare to your answer from (b)?

```{r 9c}
se_boot <- function(data, index){
  mean(Boston$medv[index])
}

set.seed(1)
(se_estimate <- boot(Boston, se_boot, 500))
```

The bootstrap standard error is very similar to the standard error from (b) (0.40 compared to 0.41).

#### (d) Based on your bootstrap estimate from (c), provide a 95% confidence interval for the mean of `medv`. Compare it to the results obtained using `t.test(Boston$medv)`.
Hint: You can approximate a 95% confidence interval using the
formula [$\hat{\mu} − 2SE(\hat{\mu}), \hat{\mu} + 2SE(\hat{\mu})$].

```{r 9d}
# recalculating se because it is not easy to get this value 
# directly from the output of boot() 
se <- sd(se_estimate$t)
(low_CI <- mu_hat - 2*se) # lower boundary of 95% CI
(up_CI <- mu_hat + 2*se) # upper boundary of 95% CI

t.test(Boston$medv)

```

The results of the two methods are the same up to one decimal place.

#### (e) Based on this data set, provide an estimate, $\hat{\mu}_{med}$, for the median value of `medv` in the population.

```{r 9e}

median(Boston$medv)

```


#### (f) We now would like to estimate the standard error of $\hat{\mu}_{med}$. Unfortunately, there is no simple formula for computing the standard error of the median. Instead, estimate the standard error of the median using the bootstrap. Comment on your findings.

```{r 9f}

med_boot <- function(data, index){
  median(Boston$medv[index])
}

set.seed(1)
boot(Boston, med_boot, 500)

```

The standard error of the median is 0.37

#### (g) Based on this data set, provide an estimate for the tenth percentile of `medv` in Boston suburbs. Call this quantity $\hat{\mu}_{0.1}$ (You can use the `quantile()` function.)

```{r 9g}
quantile(Boston$medv, 0.1)
```


#### (h) Use the bootstrap to estimate the standard error of $\hat{\mu}_{0.1}$. Comment on your findings.

```{r 9h}

mu10_boot <- function(data, index){
  quantile(Boston$medv[index], 0.1)
}

set.seed(1)
boot(Boston, mu10_boot, 500)
```

The estimated standard error is 0.52.
