---
title: "An Introduction to Statistical Learning: Chapter 5 Applied Exercises"
author: "Kylie Foster"
date: "1 May 2019"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS) # for lda, and contains a large collection of datasets
#library(GGally) # for ggpairs
library(tidyverse) # loads packages including ggplot2, dplyr
#library(skimr) # for nicer summaries using skim
#library(cowplot) # for plot_grid
#library(car) # for vif()
#library(olsrr) # for studentized residual plots
library(knitr) # for kable
#library(ggplot2)
library (class) # for knn()
library(ISLR) # data sets for the ISLR book
library(boot) # for cv.glm()
```

Still need to do Ex13

## Exercise 5

In Chapter 4, we used logistic regression to predict the probability of
default using `income` and `balance` on the `Default` data set. We will
now estimate the test error of this logistic regression model using the
validation set approach. Do not forget to set a random seed before
beginning your analysis.

#### (a) Fit a logistic regression model that uses income and balance to predict default.

```{r 5_a}

log_fit <- glm(default ~ income + balance, data = Default, family = "binomial")
summary(log_fit)
```

#### (b) Using the validation set approach, estimate the test error of this model. In order to do this, you must perform the following steps:

#### i. Split the sample set into a training set and a validation set.

```{r 5bi}

set.seed(1)
# vector of row numbers to using in training set.
train <- sample(dim(Default)[1], dim(Default)[1]*0.8, replace = FALSE, prob = NULL) # using dim[1] 
# to work out the correct length and using 80% of the data for training

```

#### ii. Fit a multiple logistic regression model using only the training observations.

```{r 5bii}
# fitting logistic model using only training data
log_fit_train <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)

```

#### iii. Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5.

```{r 5biii}
# calculating predictions using test set
log_prob <- predict(log_fit_train, Default, type = "response")[-train] # predicted probabilities for the test data

# Getting class predictions using 0.5 as the threshold:
log_pred <- as.factor(case_when(log_prob <= 0.5 ~"No",
                                log_prob > 0.5 ~ "Yes"))

```

#### iv. Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.

```{r 5biv}

mean(log_pred != Default$default[-train]) # validation set error

```

#### (c) Repeat the process in (b) three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained.

```{r}
# Using a different seed to give a different train/test split
set.seed(2)
train2 <- sample(dim(Default)[1], 0.8*dim(Default)[1])
# fitting logistic model using train2
log_fit_train2 <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train2)
# calculating predictions using test set
log_prob <- predict(log_fit_train2, Default, type = "response")[-train2]
log_pred <- as.factor(case_when(log_prob <= 0.5 ~ "No",
                                log_prob > 0.5 ~ "Yes"))
# validation test set error
mean(log_pred != Default$default[-train2])

##
# Using another different seed to give a different train/test split
set.seed(3)
train3 <- sample(dim(Default)[1], 0.8*dim(Default)[1])
# fitting logistic model using train3
log_fit_train3 <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train3)
# calculating predictions using test set
log_prob <- predict(log_fit_train3, Default, type = "response")[-train3]
log_pred <- as.factor(case_when(log_prob <= 0.5 ~ "No",
                                log_prob > 0.5 ~ "Yes"))
# validation test set error
mean(log_pred != Default$default[-train3])

```

The validation set errors are similar, low, values for the different splits (approximately 2 to 3%).

#### (d) Now consider a logistic regression model that predicts the probability of `default` using `income`, `balance`, and a dummy variable for `student`. 
Estimate the test error for this model using the validation set approach. Comment on whether or not including a
dummy variable for `student` leads to a reduction in the test error rate.

```{r 5d}

set.seed(1)
train <- sample(dim(Default)[1], 0.8*dim(Default)[1]) # 80%/20% train/test split
# gitting logistic model, including student as a predictor
log_fit_train <- glm(default ~., data = Default, family = "binomial", subset = train)
summary(log_fit_train)
# calculating predicted probabilities using test (cross validation) set
log_prob <- predict(log_fit_train, Default, type = "response")[-train]
log_pred <- as.factor(case_when(log_prob <= 0.5 ~ "No",
                                log_prob > 0.5 ~ "Yes"))
# validation test set error
mean(log_pred != Default$default[-train])

```

Using this validation set, including a dummy variable for `student` did not lead to any meaningful reducion in the test error rate.

## Exercise 6

We continue to consider the use of a logistic regression model to
predict the probability of `default` using `income` and `balance` on the
`Default` data set. In particular, we will now compute estimates for
the standard errors of the `income` and `balance` logistic regression coefficients
in two different ways: (1) using the bootstrap, and (2) using
the standard formula for computing the standard errors in the `glm()`
function. Do not forget to set a random seed before beginning your
analysis.

#### (a) Using the `summary()` and `glm()` functions, determine the estimated standard errors for the coefficients associated with `income` and `balance` in a multiple logistic regression model that uses both predictors.

```{r 6a}
# fitting multiple logistic regression model
log_multi <- glm(default ~ income + balance, data = Default, family = "binomial")
summary(log_multi) # using summary to determine the estimated standard errors for the coefficients

```

#### (b) Write a function, `boot.fn()`, that takes as input the `Default` data set as well as an index of the observations, and that outputs the coefficient estimates for `income` and `balance` in the multiple logistic regression model.

#### (c) Use the `boot()` function together with your `boot.fn()` function to estimate the standard errors of the logistic regression coefficients for `income` and `balance`.

#### (d) Comment on the estimated standard errors obtained using the `glm()` function and using your bootstrap function.

## Exercise 7

In Sections 5.3.2 and 5.3.3, we saw that the `cv.glm()` function can be
used in order to compute the LOOCV test error estimate. Alternatively,
one could compute those quantities using just the `glm()` and `predict.glm()` functions, and a for loop. You will now take this approach
in order to compute the LOOCV error for a simple logistic
regression model on the `Weekly` data set. Recall that in the context
of classification problems, the LOOCV error is given in (5.4).

#### (a) Fit a logistic regression model that predicts `Direction` using `Lag1` and `Lag2`.

```{r 7a}
# fitting multiple logistic regression model
log_Wekly <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = "binomial")
```

#### (b) Fit a logistic regression model that predicts `Direction` using `Lag1` and `Lag2` using all but the first observation.

#### (c) Use the model from (b) to predict the direction of the first observation. You can do this by predicting that the first observation will go up if P(Direction="Up"|Lag1, Lag2) > 0.5. Was this observation correctly classified?

#### (d) Write a for loop from i = 1 to i = n, where n is the number of observations in the data set, that performs each of the following steps:

#### i. Fit a logistic regression model using all but the ith observation to predict Direction using Lag1 and Lag2.

#### ii. Compute the posterior probability of the market moving up for the ith observation.

#### iii. Use the posterior probability for the ith observation in order to predict whether or not the market moves up.

#### iv. Determine whether or not an error was made in predicting the direction for the ith observation. If an error was made, then indicate this as a 1, and otherwise indicate it as a 0.

#### (e) Take the average of the n numbers obtained in (d)iv in order to obtain the LOOCV estimate for the test error. Comment on the results.

## Exercise 8

We will now perform cross-validation on a simulated data set.

#### (a) Generate a simulated data set as follows:

`set.seed(1)`
`x = rnorm(100)`
`y = x - 2*x^2 + rnorm(100)`

In this data set, what is n and what is p? Write out the model
used to generate the data in equation form.

#### (b) Create a scatterplot of `X` against `Y`. Comment on what you find.

#### (c) Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using least squares:

#### i. $Y = \beta_0 + \beta_1X + \epsilon$

#### ii. $Y = \beta_0 + \beta_1X + beta_2X^2 + \epsilon$

#### iii. $Y = \beta_0 + \beta_1X + beta_2X^2 +\beta_3X^3 + \epsilon$

#### iv. $Y = \beta_0 + \beta_1X + beta_2X^2 + \beta_3X^3 + \beta_4X^4 + \epsilon$.

Note you may find it helpful to use the data.frame() function
to create a single data set containing both `X` and `Y`.

#### (d) Repeat (c) using another random seed, and report your results. Are your results the same as what you got in (c)? Why?

#### (e) Which of the models in (c) had the smallest LOOCV error? Is this what you expected? Explain your answer.

#### (f) Comment on the statistical significance of the coefficient estimates that results from fitting each of the models in (c) using
least squares. Do these results agree with the conclusions drawn based on the cross-validation results?

## Exercise 9

We will now consider the `Boston` housing data set, from the `MASS`
library.

#### (a) Based on this data set, provide an estimate for the population mean of `medv`. Call this estimate $\hat{\mu}$.

#### (b) Provide an estimate of the standard error of ˆμ. Interpret this result.
Hint: We can compute the standard error of the sample mean by
dividing the sample standard deviation by the square root of the
number of observations.

#### (c) Now estimate the standard error of ˆμ using the bootstrap. How does this compare to your answer from (b)?

#### (d) Based on your bootstrap estimate from (c), provide a 95% confidence interval for the mean of `medv`. Compare it to the results
obtained using `t.test(Boston$medv)`.
Hint: You can approximate a 95% confidence interval using the
formula [ˆμ − 2SE(ˆμ), ˆμ + 2SE(ˆμ)].

#### (e) Based on this data set, provide an estimate, ˆμmed, for the median value of `medv` in the population.

#### (f) We now would like to estimate the standard error of ˆμmed. Unfortunately, there is no simple formula for computing the standard
error of the median. Instead, estimate the standard error of the median using the bootstrap. Comment on your findings.

#### (g) Based on this data set, provide an estimate for the tenth percentile of `medv` in Boston suburbs. Call this quantity ˆμ0.1. (You
can use the `quantile()` function.)

#### (h) Use the bootstrap to estimate the standard error of ˆμ0.1. Comment on your findings.